Category-based and location-based volitional covert attention affect memory at different timescales

### 2023-03-25 paper/main.tex
minor wording fix
added order effect analyses to results section

@@

Next, we turned to identifying potential differences in participants' behaviors
across the two experimental conditions. The main difference between the
conditions' procedures was in how long participants were asked to maintain the
same focus of category-based and location-based attention, across successive
image presentations. Therefore, differences in participants' behaviors across
the two conditions might reflect differences in the timescales of the relevant
cognitive processes. We compared participants' familiarity ratings for images
at each attention level across the two conditions. We saw no evidence that
people rated fully attended images ($t(51) = -0.649, p = 0.519$),
attended-category (but not location) images ($t(51) = 1.163, p = 0.250$),
unattended category and location images ($t(51) = -1.600, p = 0.116$), or novel
images ($t(51) = -0.435, p = 0.665$) differently across the two conditions.
However, participants in the variable attention condition rated
attended-location (but unattended category) images as more familiar than
participants in the sustained attention condition ($t(51) = 2.174, p = 0.034$).
This suggests that the timescales over which location-based versus
category-based attention tend to operate may differ.

\textbf{JRM STOPPED HERE}

@@

\begin{figure*}[tp]
  \centering \includegraphics[width=0.75\textwidth]{figures/discussion_schematic}

  \caption{\textbf{Timecourse of encoding and retrieval effects of
  feature-based and location-based attention.} The focus of location-based
  attention (green) may be modulated quickly and rapidly enhances memory
  encoding for nearby stimuli. The focus of feature-based attention (purple) is
  modulated more slowly, and serves to suppress unattended stimulus features.
  The effects of location-based attention on memory persist for longer than the
  effects of feature-based attention. Sustained attention (darker shading)
  yields more robust enhancement and suppression than shorter-term (variable)
  attention (light shading). The vertical gray line on the left denotes the
  duration of a single stimulus presentation (the upper bound by which
  location-based attention begins to affect encoding, and the lower-bound by
  which feature-based attention begins to affect encoding). The vertical gray
  line on the right separates encoding from immediate subsequent retrieval of
  the encoded information.}

\label{fig:model}
\end{figure*}

>>>>

\begin{figure*}[tp]
	\centering
	\includegraphics[width=0.7\textwidth]{figs/familiarity_by_attention_level}

  \caption{\textbf{Familiarity by attention level.} The bars display the
  average familiarity ratings participants gave to images from the same
  category and location as the attention cue (fully attended), the same
  category (but opposite location) as the attention cue (attended category),
  the same location (but opposite category) as the attention cue (attended
  location), the opposite category and location as the attention cue
  (unattended), or novel images. The left panel displays familiarity ratings
  from the sustained attention condition and the right panel displays
  familiarity ratings from the variable attention condition. All error bars
  denote across-participant bootstrap-estimated 95\% confidence intervals. For results
  sub-divided by stimulus category, see Figure~\ratingsByCategory.}

  \label{fig:familiarity}
  \end{figure*}

@@

Next, we turned to identifying potential differences in participants' behaviors
across the two experimental conditions. The main difference between the
conditions' procedures was in how long participants were asked to maintain the
same focus of category-based and location-based attention, across successive
image presentations. Therefore, differences in participants' behaviors across
the two conditions might reflect differences in the timescales of the relevant
cognitive processes. We compared participants' familiarity ratings for images
at each attention level across the two conditions. We saw no evidence that
people rated fully attended images ($t(51) = -0.649, p = 0.519$),
attended-category (but not location) images ($t(51) = 1.163, p = 0.250$),
unattended category and location images ($t(51) = -1.600, p = 0.116$), or novel
images ($t(51) = -0.435, p = 0.665$) differently across the two conditions.
However, participants in the variable attention condition rated
attended-location (but unattended category) images as more familiar than
participants in the sustained attention condition ($t(51) = 2.174, p = 0.034$).
This suggests that category-based versus location-based attention may operate
over different timescales.

\begin{figure*}[tp]
	\centering
	\includegraphics[width=0.7\textwidth]{figs/encoding_and_retrieval_timecourses}

  \caption{\textbf{Familiarity ratings by serial positions and attention
  level.} \textbf{A. Subsequent familiarity ratings by presentation position
  (sustained attention condition).} The curves' colors denote the attention
  levels of each presented image. The $x$-axis denotes the presentation
  positions of each image within the sequence of 10 composite image pairs
  during the run when it was presented. The $y$-axis denotes the average
  familiarity ratings later given to the corresponding items. \textbf{B.
  Subsequent familiarity ratings by presentation position (variable attention
  condition).} This panel is in the same format as Panel A, but displays
  ratings for the variable attention condition. \textbf{C. Familiarity ratings
  by memory probe position (sustained attention condition).} The curve's colors
  denote the attention levels (or novelty) of each probe image. The $x$-axis
  denotes the position of each probed image within the sequence of 40 images
  that participants judged during the memory phase of the experiemnt. The
  $y$-axis denotes the average familiarity ratings given to the corresponding
  probes. \textbf{D. Familiarity ratings by memory probe position (variable
  attention condition).} This panel is in the same format as Panel C, but
  displays ratings for the variable attention condition. All panels: error
  ribbons denote across-participant bootstrap-estimated 95\% confidence
  intervals.}

  \label{fig:timecourses}
  \end{figure*}

Given the above results suggesting potential differences in the timescales of
category-based and location-based attention, we carried out two additional
exploratory analyses aimed at identifying other timing effects. First, we
wondered whether participants' familiarity ratings might show serial position
effects analogous to those reported in classic recognition memory
studies~\citep[e.g.,][]{Neat93b, McElDosh89, WickNorm66}. For each participant,
for each composite image pair (presented in each trial during the study phase
of the experiment), we labeled each composite's face and place image component
according to whether it matched the cued category and/or location. We discarded
any face or place images that did not appear in the participants' memory phase.
We tagged the remaining (probed) images with the familiarity ratings that
participants would later give the images during the memory phase and plotted
these ratings against the images' presentation positions
(Figs.~\ref{fig:timecourses}A and B). Across both experimental conditions and
across all serial positions, we generally found that the average ordering of
familiarity ratings by attention level (Fig.~\ref{fig:familiarity}) were
preserved. This suggests that the encoding-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the study phase of the
experiment). Second, we carried out an analogous analysis to identify potential
serial position effects of recall order. For each probed item a participant
rated during the memory phase, we assigned the image a label according to
whether the participant's attention cue (at the time the image was presented as
part of its composite pair) matched the image's category and/or location, or
whether the image was novel (i.e., not presented during the study phase).
Again, we found that (across both experimental conditions and all probe
positions) in general the average ordering of familiarity ratings by attention
level (Fig.~\ref{fig:familiarity}) were preserved. This suggests that the
retrieval-related affects of attention on subsequent recognition memory are
relatively stable over time (e.g., we did not observe clear primacy or recency
effects during the memory phase of the experiment).

\textbf{JRM STOPPED HERE}
%note: discuss response bias analyses next (also add to supplement)

@@

% \begin{figure*}[tp]
%   \centering \includegraphics[width=0.75\textwidth]{figures/discussion_schematic}

%   \caption{\textbf{Timecourse of encoding and retrieval effects of
%   feature-based and location-based attention.} The focus of location-based
%   attention (green) may be modulated quickly and rapidly enhances memory
%   encoding for nearby stimuli. The focus of feature-based attention (purple) is
%   modulated more slowly, and serves to suppress unattended stimulus features.
%   The effects of location-based attention on memory persist for longer than the
%   effects of feature-based attention. Sustained attention (darker shading)
%   yields more robust enhancement and suppression than shorter-term (variable)
%   attention (light shading). The vertical gray line on the left denotes the
%   duration of a single stimulus presentation (the upper bound by which
%   location-based attention begins to affect encoding, and the lower-bound by
%   which feature-based attention begins to affect encoding). The vertical gray
%   line on the right separates encoding from immediate subsequent retrieval of
%   the encoded information.}

% \label{fig:model}
% \end{figure*}
